---
title: "Coordinates Assessment"
author: Tyler Brown
date: 2017-11-28
output:
  pdf_document:
    highlight: zenburn
---
```{r, include = FALSE}
library(dplyr)
library(knitr)
library(readr)
```

```{r, include = FALSE}
# Data in question
fixed_loc <- read_csv("../data/fixed_loc.csv")
combined <- read_csv("../data/Combined_data.csv")
```

We need to verify that imputed coordinates are making sense. We start
with the "combined" data which includes mostly unique `pid-year`
combinations.

Description of Datasets and Imputations
---------------------------------------

```{r, echo = FALSE}
kable(tibble( row_count = nrow(combined),
             col_count = ncol(combined)),
      caption = "'Combined' dataset dimensions")
```

Given these dimensions, the `Latitude` and `Longitude` columns are
both filled out about `65%` of the time.

```{r, echo = FALSE}
td <- table(complete.cases(combined %>% select( Latitude, Longitude )))
per <- round(td['FALSE'] / (td['TRUE'] + td['FALSE']), 4) * 100

kable(tibble( complete = td['TRUE'],
             incomplete = td['FALSE'],
             percent_incomplete = per,
             total = td['TRUE'] + td['FALSE']),
      caption = "Coordinate Pair Completion Rate")
```

```{r, echo = FALSE}

unique_coordinates <- combined %>% filter(
                                !complete.cases(combined %>%
                                select( Latitude, Longitude ))
                                )
#dim(unique_coordinates)

uc_pid_grp <- unique_coordinates %>%
    group_by(PID) %>%
    count()

kable(tibble( incomplete = nrow(unique_coordinates),
             PID = nrow(uc_pid_grp) ),
      caption = "Count of unique PID with Incomplete Coordinate Pairs")
```


The second file, `fixed_loc.csv`, includes imputed values for the missing
coordinate pairs. We imputed coordinates from [openaddresses.io](https://openaddresses.io/) for the missing data. The number of imputed values
do not account for all of the incomplete cases.

```{r, echo = FALSE}
table3_PID <- nrow(uc_pid_grp)
row_count <- nrow(fixed_loc)
percent_imputed <- round(row_count/table3_PID, 4) * 100

kable(tibble(table3_PID = table3_PID,
             fixed_loc_row_count = row_count,
             percent_imputed = percent_imputed),
      caption = "Percent Imputed by `Fixed Locations`")
```

Quality of Imputed Coordinates
------------------------------

Let's take a closer look at `fixed_loc.csv` to see if coordinate
pairs were successfully imputed.

```{r, echo = FALSE}
f1 <- fixed_loc %>%
    group_by(PID) %>%
    count() %>%
    arrange(desc(n))

per_pid <- round(nrow(f1) / nrow(fixed_loc), 4) * 100

kable(tibble(row_count = nrow(fixed_loc),
             unique_PID = nrow(f1),
             percent_unique_pid = per_pid),
      caption = "Check: Is each row a unique PID?")
```

The number of unique `PID` values is almost double. This is most likely
caused by a `one-to-many` join when a `one-to-one` join was intended.


```{r, echo = FALSE}
f2 <- fixed_loc %>%
    group_by(PID, LAT, LON) %>%
    count() %>%
    group_by(n) %>%
    summarize(
        unique_row_count = n()
    ) %>%
    arrange(desc(unique_row_count))

kable(as_tibble(head(f2, n = 10)),
      caption = "Top 10 Multiple PID-Coordinate Combinations")

summary(f2$n)
```

The `fixed_loc.csv` file clearly needs some work. This work will need
to be done in a later version of the application. Let's try to put
together a minimal version of `fixed_loc.csv` that may be usable for
an initial imputation.

We have multiple coordinate pairs matched to `PID` values. I'm going
to group by `PID` and take the first match.

```{r, echo = FALSE}
first_match <- fixed_loc %>%
    mutate(
        coordinate_pair = paste(LAT, LON, sep=";")
    ) %>%
    group_by(PID) %>%
    summarize(
        cpair = first(coordinate_pair),
        lat = strsplit(cpair, ";")[[1]][1],
        lng = strsplit(cpair, ";")[[1]][2]
    )

kable(tibble(
    first_match_rows = nrow(first_match),
    table5_row_count = nrow(fixed_loc)),
    caption = "Comparing First Group Row Count to Table 5")
```

Now that we have a file without duplicate `PID` values, let's see how
unique imputed coordinates are compared to the `Combined` file.

```{r, echo = FALSE}
fmatch_agg <- first_match %>%
    group_by(cpair) %>% count() %>%
    group_by(n) %>%
    summarize(
        coordinate_pair_count = n()
    ) %>%
    arrange(desc(coordinate_pair_count))

print("Quantiles of an Aggregate on the `first_match` table")
quantile(fmatch_agg$coordinate_pair_count)
```

We can clearly see here that some `PID` values will share a coordinate
pair. This is to be expected on some level because we ignored the unit
number when matching coordinates. It wouldn't make sense for the unit
number within a property to have a meaningfully different coordinate
value.

The last thing to do for this version is to go through and see if
a coordinate pair with multiple `PID` values is substantively valid.

```{r, echo = FALSE}
cfirst <- combined %>%
    rowwise() %>%
    mutate(
        st_address = paste(as.character(ST_NUM),
                           as.character(ST_NAME),
                           as.character(ST_NAME_SUF),
                           #as.character(UNIT_NUM),
                           as.character(ZIPCODE))
    )

cfirst_grp <- cfirst %>%
    group_by(PID, st_address) %>%
    count() %>%
    ungroup()

imputed_coords <- inner_join(x = cfirst_grp, y = first_match, by = "PID")


kable(tibble(
    unique_pid_addr_in_combined = nrow(cfirst_grp),
    imputed_coord_count = nrow(imputed_coords)),
    caption = "Number of unique addresses imputed.")
```

Now we need to see if the number of coordinates match up with the
number of units for each address. A `PID` will correspond to each
unit within a property.

```{r, echo = FALSE}
c1 <- imputed_coords %>%
    group_by(st_address, cpair) %>%
    count()

c2 <- imputed_coords %>%
    group_by(st_address) %>%
    count()

c3 <- imputed_coords %>%
    group_by(cpair) %>%
    count()

kable(tibble(
    st_address_cpair_rows = nrow(c1),
    st_address_rows = nrow(c2),
    cpair_rows = nrow(c3),
    st_addr_cpair_to_cpair_ratio = nrow(c1)/nrow(c3)),
    caption = "Granularity of Coordinates")
```


In the above table, it's clear to see that the granularity of imputed
coordinates is an order of magnitude less than the granularity provided
by each street address.


Comparing Non-imputed Granularity of Coordinates
-------------------------------------------------

Let's see how granular non-imputed coordinates are for the city of
Boston.

```{r, echo = FALSE, warning = FALSE}

cgiven <- cfirst[complete.cases(cfirst %>% select(Longitude, Latitude)),]
cgiven <- cgiven %>%
    rowwise() %>%
    mutate(
        cpair = paste(as.character(Longitude),
                      as.character(Latitude), sep = ";")
    )

cgiven_grp <- cgiven %>%
    group_by(PID, st_address, cpair) %>%
    count() %>%
    ungroup()

cg1 <- cgiven %>%
    group_by(st_address, cpair) %>%
    count()

cg2 <- cgiven %>%
    group_by(st_address) %>%
    count()

cg3 <- cgiven %>%
    group_by(cpair) %>%
    count()

kable(tibble(
    st_address_cpair_rows = nrow(cg1),
    st_address_rows = nrow(cg2),
    cpair_rows = nrow(cg3),
    st_addr_cpair_to_cpair_ratio = nrow(cg1)/nrow(cg3)),
    caption = "Granularity of Non-Imputed Coordinates for Combined")
```

It appears that imputed coordinates are (`6.5:1`)considerably less 
granular than the given coordinates (`4:1`).

Summary
-------

* The initial `fixed_loc` file had some issues with multiple
  coordinates being matched to a `PID` in several cases.
* As a first version fix, I grouped on each `PID` within `fixed_loc`
  and kept the first coordinate pair.
* The ratio of `PID` address and coordinate pair over `PID`
  coordinate pair is similar for the imputed and non-imputed
  coordinates.


To Do
-----

* Given the `6.5:1` for imputed and `4:1` for non-imputed ratio of street
  addresses to unique coordinate pairs for both imputed and non-imputed 
  values, we will probably need to reconsider our mapping strategy.
* When I comment out `UNIT_NUM`, the ratios drop to `1.3:1` for
  imputed and `3:1` for non-imputed. It seems as though the imputed
  values may be more granular from this perspective.
