---
title: "Boston Property Assessment Data Audit 2014-2017"
author: Tyler Brown
date: 2017-10-13
output:
  pdf_document:
    highlight: zenburn
---
```{r, include = FALSE}
library(dplyr)
library(ggplot2)
library(readr)
library(scales)
library(stringr)
library(tidyr)
library(knitr)

# custom
source("data_audit.R")
```

```{r, include = FALSE}
# Boston property assessment data
combined <- read_csv("../data/Combined_data.csv")
```

Summary
=======

The purpose of a data audit is to dig up skeletons in the closet. Let's
review the skeletons we found:

* `PID-year` is our unique ID for this dataset. It occurs more than once
  less than one percent of the time.
* About 27% of the variables are either filled out more than 70% of the
  time or failed a count of blank values test.
* Location appears to be most easily mapped using `ZIPCODE` instead of
  Latitude and Longitude because they're missing about 35% of the time.
* `GROSS_TAX` appears to be filled out without any incorrect entries.
* Additional data audit information can be found in the 
  "property-assessment-2014-2017" pdf.


Useful Columns
--------------

We can get an initial idea about which columns are useful by seeing
how often they're filled out. The following columns are not blank
or NA more than 70% of the time.

```{r, echo = FALSE}
results2 <- results %>%
    filter(percent_nan <= 30 | is.na(percent_nan)) %>%
    arrange(variables)
kable(results2, caption = "Percentage of columns at least 70% filled out")
```

Most of our variables are missing from the above table. The next
table shows the percentage of variables which are at least 70% populated.

```{r, echo = FALSE}
kable(tibble(
    populated_70 = round(length(results2$variables) /
                         length(results$variables), 4) * 100
), caption = "Percentage of Variables 70% populated including NAs")
```

Next we want to double check those columns which came up as `NA` for
the 'is blank' tests. 

```{r, echo = FALSE}
results3 <- results %>%
    filter(is.na(num_blanks) & is.na(num_nans))

kable(results3, caption = "NA values for Is Blank Test")
```

Come back to this later.

Checking for each Unique ID in Each Row
---------------------------------------

Concatenated `PID` and `year`. This should be a unique value for each
row. 

```{r, echo = FALSE}
uid <- combined %>%
    mutate( yearPID = paste(Year, PID, sep = "-") ) %>%
    group_by(yearPID) %>%
    count() %>%
    group_by(n) %>%
    count()

kable(tibble(non_unique_ids = sum(uid[uid$n >= 2,]$nn) / sum(uid$nn)),
      caption = "Percentage of unique values occuring more than once.")
```

It looks like almost all of the time we have unique IDs. However, we will
want to filter out the `PID-year` values which are not unique.

Checking for Variables that can be used to figure out location
--------------------------------------------------------------

Since a nontrivial amount of our project involves mapping, we want to
make sure we're able to map `PID-year` to a location. In the appendix,
`Latitude` and `Longitude` are shown to have `NA` values about 35% of the
time. Maybe we can find a way to impute coordinate values if it makes
sense.

The `ZIPCODE` variable appears to almost be completely filled out. This
doesn't really matter for location purposes if all the zip codes are
identical. Let's see how much fidelity we're getting out of `ZIPCODE`.

```{r, echo = FALSE}
zipp <- combined %>%
    group_by(ZIPCODE) %>%
    count() %>%
    arrange(desc(n))

kable(zipp, caption = "Zip Code by Frequency")
```

It seems like trying the zip code route for location may be the best
approach given currently available data.

Checking variables that can be used to measure value.
-----------------------------------------------------

```{r, echo = FALSE}
divisor <- 100000000
grosstax <- combined %>%
    mutate(GROSS_TAX_BIN = floor(GROSS_TAX / divisor)) %>%
    group_by(GROSS_TAX_BIN) %>%
    count()

kable(grosstax, caption = "Gross Tax Frequency, $100 Million Bins")
```

There doesn't seem to be any junk data in `GROSS_TAX`.



Appendix
--------

Here's the counts of what's blank or not.

```{r, echo = FALSE}
results <- is_blank(combined) %>%
    mutate(
        percent_nan = round(num_nans / dim(combined)[1], 4) * 100
    ) %>%
    arrange(variables)

kable(results, caption = "Count of blanks and NAs")
```

